PANCHANGE_data_merging

### 2020-08-17 PANCHANGE_data_merging.Rmd
created merging script

---
title: "PANCHANGE_data_preprocessing"
author: "Katie Thompson, Topher Huebel, Molly Davies, Kirstin Purves"
date: Sys.Date()
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes

params:
  glad.dem.raw: edgi.dem.raw

---

@@

# Specify cohort
```{r Specify cohort}
GLAD = TRUE
EDGI = FALSE
NBR = FALSE
RAMP = FALSE
```

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("eeptools")
#install.packages("skimr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(eeptools) # Needed to calculate age from birthdate
library(tidyverse)
```

# Colour palettes
Define colours for plotting this are the standard coping colours
```{r Colour palettes: COPING}
COPINGpalette2 <- c("#78D9C5",
                    "#F5BE5E")

COPINGpalette3 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9")

COPINGpalette4 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73")

COPINGpalette5 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73", 
                    "#FFED98")

COPINGpalette6 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73",
                    "#FFED98",
                    "#BFD2EB")

COPINGpalette7 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73", 
                    "#FFED98",
                    "#BFD2EB", 
                    "#808080")

COPINGpaletteGRAD <- c("#F5BE5E",
                       "#FFD284",
                       "#FFEED1",
                       "#B5B5B5",
                       "#DEFFF8",
                       "#94F6E1",
                       "#78D9C5")

COPINGNeuCenterpalette <- c("#78D9C5",
                            "#808080",
                            "#F5BE5E")

RAMPworseGRADpalette <- c("#78D9C5",
                          "#FFEED1",
                          "#F5BE5E",
                          "#FFB1B5")
GLADpalette = c("#efc00b", 
                "#b7dee8")  
```

Choose in this chunk which palette to use
04.07.2020 - Default to 2 colour COPING palette
```{r Choose colour palette}
palette = COPINGpalette2
```

#Add_numeric function
Function used to convert character variables into numeric variables. 
```{r add_numeric function}
add_numeric <- function(dat, exclude = NULL) {
  dat_fct <- sjlabelled::as_label(dat)

  dat <- dat[!colnames(dat) %in% exclude]
  colnames(dat) <- paste(colnames(dat), "numeric", sep = "_")
  return(bind_cols(dat_fct, dat))
}
```

# Data import and merging

Note: GLAD only receives current measures and RAMP questionnaire
Note: All COPING surveys are separate surveys

@@

# Demographics
## GLAD
### Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.dem.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.dem.raw)
colnames(glad.dem.raw)
}
```

### Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.prepandemic",
                  "Age_uncleaned",
                  "EduYrs")

glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate.prepandemic = startDate, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.dem.raw.id)
# Inspect colnames
colnames(glad.dem.raw.id)
#Differences
dim(glad.dem.raw)[1]-dim(glad.dem.raw.id)[1]
}
```

```{r glad copy dem}
if(GLAD == TRUE){
dem.raw.id <- glad.dem.raw.id
}
```

```{r glad dem summary}
if(GLAD == TRUE) {
summarytools::dfSummary(
  dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

## EDGI
### Read in data
```{r edgi dem cols}
if(EDGI == TRUE){
  edgi.dem.raw <- readRDS(file = paste0(data.raw_path, "/edgi/dem_edgi.rds"))
  dim(edgi.dem.raw)
  colnames(edgi.dem.raw)
}
```

### Select variables
```{r edgi dem cleaning}
if(EDGI == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.prepandemic",
                        "Age_uncleaned",
                        "EduYrs")

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(edgi.dem.raw.id)
  # Inspect colnames
  colnames(edgi.dem.raw.id)
  #Differences
  dim(edgi.dem.raw)[1]-dim(edgi.dem.raw.id)[1]
}
```

```{r edgi copy dem}
if(EDGI == TRUE){
  dem.raw.id <- edgi.dem.raw.id
}
```

```{r edgi dem summary}
if(EDGI == TRUE) {
summarytools::dfSummary(
  dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

## NBR
### Read in data
```{r nbr dem cols}
if(NBR == TRUE){
  nbr.dem.raw <- readRDS(file = paste0(data.raw_path, "/nbr/dem_nbr.rds"))
  dim(nbr.dem.raw)
  colnames(nbr.dem.raw)
}
```

### Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "Age_uncleaned",
                        "EduYrs")

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, # Date when participant signed up to nbr (is in the coping gad chunk)
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear_unc = demographics.year, # Birthyear
      Birthmonth = demographics.month, #Birthmonth
      Birthday = demographics.day, #Birthday
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
#      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(nbr.dem.raw.id)
  # Inspect colnames
  colnames(nbr.dem.raw.id)
  #Differences
  dim(nbr.dem.raw)[1]-dim(nbr.dem.raw.id)[1]
}
```

```{r nbr dem calculating age from birthday}
if(NBR == TRUE) {
#add 1900 to birthyear
nbr.dem.raw.id$Birthyear_numeric <- nbr.dem.raw.id$Birthyear_unc_numeric + 1900

#create dob as date format
nbr.dem.raw.id$Birthdate <- as.Date(paste(nbr.dem.raw.id$Birthyear_numeric,
                                  nbr.dem.raw.id$Birthmonth_numeric,
                                  nbr.dem.raw.id$Birthday_numeric,
                                  sep = "-"))

head(nbr.dem.raw.id$Birthdate)

#Calculate dem.age(remove NAs as it throws an error)
ages <- age_calc(na.omit(nbr.dem.raw.id$Birthdate), enddate = Sys.Date(), units = "years")

#Create dem.agecolumn
nbr.dem.raw.id$Age_uncleaned<- NA

#Populate with values (excluding NAs)
nbr.dem.raw.id$Age_uncleaned[!is.na(nbr.dem.raw.id$Birthdate)] <- ages

#Round age down
nbr.dem.raw.id$Age_uncleaned <- floor(nbr.dem.raw.id$Age_uncleaned)

#Visualise
head(nbr.dem.raw.id$Age_uncleaned)
head(nbr.dem.raw.id$Birthdate)
}
```

### Cohort information
Information on INTERVAL, COMPARE, and STRIDES
http://www.donorhealth-btru.nihr.ac.uk/studies/
```{r nbr cohort information}
if(NBR == TRUE){
  nbr.cohort.raw <- readxl::read_xlsx(
    path = paste0(data.raw_path, "/nbr/coping_participants_export_20200730.xlsx")
    )

  dim(nbr.cohort.raw)
  colnames(nbr.cohort.raw)
  }
```

```{r nbr cohort cleaning}
if(NBR == TRUE) {
exclude_cols_cohort_nbr <- c("ID",
                  "Sample",
                  "Age_uncleaned",
                  "EduYrs")

nbr.cohort.raw.id <- nbr.cohort.raw %>%
  drop_na(`NBR ID`) %>% # Drop NAs
  distinct(`NBR ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`NBR ID`, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         Cohort = cohort, #cohort information
#         startDate.coping, #Date when participant signed up to NBR COPING (is in the coping gad chunk)
#         Age_uncleaned = age, # Age
#         Gender = gender, # Gender
#         Sex = demographics.select_questionnaire_items_medical, #Sex
#         EduYrs = demographics.school_education_school_years.txt, # Years in education
#         demographics.college_or_university_degree, # level of education - univeristy
#         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#         demographics.cses_or_equivalent, # level of education - cse
#         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity.unc = ethnicity # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_cohort_nbr) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.cohort.raw.id)
# Inspect colnames
colnames(nbr.cohort.raw.id)
#Differences
dim(nbr.cohort.raw)[1]-dim(nbr.cohort.raw.id)[1]
}
```

```{r nbr cohort recode ethnicity}
if(NBR == TRUE) {
#Recode details ethnicity info
nbr.cohort.raw.id$Ethnicity_full <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White - British",
                                             "B" = "White - Irish",
                                             "C" = "White - Other",
                                             "D" = "Mixed - White & Black Caribbean",
                                             "E" = "Mixed - White & Black African",
                                             "F" = "Mixed - White & Asian",
                                             "G" = "Mixed - Other",
                                             "H" = "Asian - Indian",
                                             "J" = "Asian - Pakistani",
                                             "K" = "Asian - Bangladeshi",
                                             "L" = "Asian - Other",
                                             "M" = "Black - Caribbean",
                                             "N" = "Black - African",
                                             "P" = "Black - Other",
                                             "R" = "Chinese - Chinese",
                                             "S" = "Chinese - Other",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)

#Match GLAD/EDGI/RAMP categories
nbr.cohort.raw.id$Ethnicity <- recode_factor(nbr.cohort.raw.id$Ethnicity.unc,
                                             "A" = "White, white European or Caucasian",
                                             "B" = "White, white European or Caucasian",
                                             "C" = "White, white European or Caucasian",
                                             "D" = "Mixed or multiple ethnic origins",
                                             "E" = "Mixed or multiple ethnic origins",
                                             "F" = "Mixed or multiple ethnic origins",
                                             "G" = "Mixed or multiple ethnic origins",
                                             "H" = "Asian or Asian British (including Chinese)",
                                             "J" = "Asian or Asian British (including Chinese)",
                                             "K" = "Asian or Asian British (including Chinese)",
                                             "L" = "Asian or Asian British (including Chinese)",
                                             "M" = "Black or Black British",
                                             "N" = "Black or Black British",
                                             "P" = "Black or Black British",
                                             "R" = "Asian or Asian British (including Chinese)",
                                             "S" = "Asian or Asian British (including Chinese)",
                                             "W" = NULL, ## Asking NBR what this level means
                                             "Z" = NULL,
                                             "99" = NULL)
}
```

```{r nbr combine demographics}
if(NBR == TRUE){
dem.raw.id <- left_join(nbr.dem.raw.id, nbr.cohort.raw.id, by = c("Sample", "ID"))

# Inspect dimensions
dim(dem.raw.id)
# Inspect colnames
colnames(dem.raw.id)
#Differences
dim(nbr.dem.raw.id)[1]-dim(dem.raw.id)[1]
}
```

```{r nbr dem summary}
if(NBR == TRUE) {
summarytools::dfSummary(
  dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

## RAMP
### Read in data
```{r ramp dem cols}
if(RAMP == TRUE){
  ramp.dem.raw <- readRDS(file = paste0(data.raw_path, "/ramp/dem_ramp.rds"))
  dim(ramp.dem.raw)
  colnames(ramp.dem.raw)
}
```

### Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample")

  ramp.dem.raw.id <- ramp.dem.raw %>%
    drop_na(`Login ID`) %>% # Drop NAs
    distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, #Date when participant signed up to ramp  (is in the coping gad chunk)
      age_category_full = demographics.how_old_are_you, # Age
      Gender = demographics.which_gender_do_you_identify_with, # Gender (what [GLAD] vs which [RAMP])
#     EduYrs = demographics.university_education_attend_school.txt, # Years in education - not in RAMP dataset
#     demographics.college_or_university_degree, # level of education - univeristy - level of education is in employment file
#     demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
#     demographics.o_levelsgcses_or_equivalent, # level of education - gcse
#     demographics.cses_or_equivalent, # level of education - cse
#     demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Transgender_uncleaned = demographics.do_you_identify_as_transgender,
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(ramp.dem.raw.id)
  # Inspect colnames
  colnames(ramp.dem.raw.id)
  #Differences
  dim(ramp.dem.raw)[1]-dim(ramp.dem.raw.id)[1]
}
```

```{r ramp copy dem}
if(RAMP == TRUE){
  dem.raw.id <- ramp.dem.raw.id
}
```

# Demographics variable creation
## Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16

# identify number of outliers
length(
  which(
  dem.raw.id$Age_uncleaned > age_upper_limit |
    dem.raw.id$Age_uncleaned < age_lower_limit
     )
)

# remove age outliers
dem.raw.id <- dem.raw.id %>%
  mutate(
    Age = 
      if_else(
        Age_uncleaned > age_upper_limit | 
        Age_uncleaned < age_lower_limit, 
        true = NA_real_,
        false = Age_uncleaned,
        missing = NA_real_
      )
  )

# Inspect variable
dem.raw.id %>%
  descr(
    Age,
    #stats = "common"
    )
}
```

## Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ 1,
          Age >= 19 & Age <= 25 ~ 2,
          Age >= 26 & Age <= 35 ~ 3,
          Age >= 36 & Age <= 45 ~ 4,
          Age >= 46 & Age <= 55 ~ 5,
          Age >= 56 & Age <= 65 ~ 6,
          Age >= 66 & Age <= 70 ~ 7,
          Age >= 71 & Age <= 75 ~ 8,
          Age >= 76 & Age <= 80 ~ 9,
          Age >= 81 & Age <= 85 ~ 10,
          Age >= 86 & Age <= 90 ~ 11,
          Age >= 91 & Age <= 100 ~ 12,
          Age >= 101 & Age <= 120 ~ 13 # oldest person in the world is 117 years
                )
      )

dem.raw.id <- dem.raw.id %>%
    mutate(
      age_category_full =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

dem.raw.id %>%
  freq(age_category_full)
}
```

## Highest education (prepandemic, GLAD & EDGI)

Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education prepandemic}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

#recode the numeric version into a factor
dem.raw.id <- dem.raw.id %>%
  mutate(
    highest_education_prepan =
      recode_factor(
        highest_education_prepan_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

dem.raw.id %>%
  freq(highest_education_prepan)
}
```

Check highest education coding
```{r Highest education prepandemic check}
if(GLAD == TRUE | EDGI == TRUE) {
dem.raw.id %>%
  select(highest_education_prepan,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```

```{r dem summary}
if(GLAD == TRUE | EDGI == TRUE) {
summarytools::dfSummary(
  dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
}
```

# Employment

## GLAD
### Read in data
```{r glad employment cols}
if(GLAD == TRUE){
glad.employment.raw <- readRDS(file = paste0(data.raw_path, "/glad/employment_coping_glad.rds"))
dim(glad.employment.raw)
colnames(glad.employment.raw)
}
```

### Select variables
```{r glad employment cleaning}
if(GLAD == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

glad.employment.raw.id <- glad.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.employment.raw.id)
# Inspect colnames
colnames(glad.employment.raw.id)
#Differences
dim(glad.employment.raw)[1]-dim(glad.employment.raw.id)[1]
}
```

```{r glad copy employment}
if(GLAD == TRUE){
  employment.raw.id <- glad.employment.raw.id
}
```

## EDGI
### Read in data
```{r edgi employment cols}
if(EDGI == TRUE){
edgi.employment.raw <- readRDS(file = paste0(data.raw_path, "/edgi/employment_coping_edgi.rds"))
dim(edgi.employment.raw)
colnames(edgi.employment.raw)
}
```

### Select variables
```{r edgi employment cleaning}
if(EDGI == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

edgi.employment.raw.id <- edgi.employment.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(
    ID = as.numeric(ID)
    ) %>%
  select(
         Sample, # Sample
         ID, # ID
         employment.phd, # level of education - univeristy
         employment.masters_degree_or_equivalent, # level of education - univeristy
         employment.postgraduate_degree_or_equivalent, # level of education - univeristy
         employment.bachelors_degree_or_equivalent, # level of education - univeristy
         employment.a_levelsas_levels_or_equivalent, # level of education - alevel
         employment.o_levelsgcses_or_equivalent, # level of education - gcse
         employment.cses_or_equivalent, # level of education - cse
         employment.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         employment.government_work_key_workers # key worker status
         ) %>%
  add_numeric(., exclude = exclude_cols_employment) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.employment.raw.id)
# Inspect colnames
colnames(edgi.employment.raw.id)
#Differences
dim(edgi.employment.raw)[1]-dim(edgi.employment.raw.id)[1]
}
```

```{r edgi copy employment}
if(EDGI == TRUE){
  employment.raw.id <- edgi.employment.raw.id
}
```

## NBR
### Read in data
```{r nbr employment cols}
if(NBR == TRUE){
nbr.employment.raw <- readRDS(file = paste0(data.raw_path, "/nbr/employment_coping_nbr.rds"))
dim(nbr.employment.raw)
colnames(nbr.employment.raw)
}
```

### Select variables
```{r nbr employment cleaning}
if(NBR == TRUE){
exclude_cols_employment <- c("ID",
                  "Sample")

>>>>

---
title: "PANCHANGE_data_merging"
author: "Christopher Huebel"
date: "17/08/2020"
output: html_document
---

@@

Load packages
```{r}
library(tidyverse)
library(skimr)
```

@@

# Import data
```{r export data}
glad <- read_rds(path = paste0(data_path, "/glad.rds"))

edgi <- read_rds(path = paste0(data_path, "/edgi.rds"))

nbr <- read_rds(path = paste0(data_path, "/nbr.rds"))

ramp <- read_rds(path = paste0(data_path, "/ramp.rds"))
```

glad/edgi data frame list
```{r glad/edgi data frame list}
dfs.list <- list(
  glad,
  edgi,
  nbr,
  ramp
)
```

```{r Join all questionnaires}
dat <- plyr::join_all(
  dfs.list,
  by = c("ID", "Sample") # Alternatively you can join by several columns EDGI -> "Age", "Sex", "Birthyear"
  )

#look at the data
skimr::skim(dat)
```

```{r Save four cohorts as rds file}
saveRDS(object = dat, file = paste0(data_path, "/four_cohorts.rds"))
```
